---
title: "Reproducible Research: Peer Assessment 1"
output:
  html_document:
    keep_md: true
---


## Loading and preprocessing the data
```{r}
library(data.table)
data <- fread("unzip -cq activity.zip")
data[,date:=as.Date(date)]
```
Converted the date column to Date data type




## What is mean total number of steps taken per day?
Calculate the total number of steps per day
```{r}
stepsperday <- data[, .(steps=sum(steps, na.rm=TRUE)), by=date]
```

Make a histogram of the total number of steps taken each day
```{r}
with(stepsperday, hist(steps))
```

Calculate and report the mean and median of the total number of steps taken per day
```{r}
summary(stepsperday$steps)[c("Mean", "Median")]
```




## What is the average daily activity pattern?
Find the number of steps taken during intervals averaged across all days
```{r}
stepsperinterval <- data[, .(steps=mean(steps, na.rm=TRUE)), by=interval]
```

Make a time series plot of the 5-minute interval and the average number of steps taken
```{r}
with(stepsperinterval, plot(interval, steps, type="l"))
```

Which 5-minute interval contains the maximum number of steps?
```{r}
stepsperinterval[which.max(steps),interval]
```




## Imputing missing values
Calculate and report the total number of missing values in the dataset
```{r}
missingsteps <- data[, .I[is.na(steps)]]
length(missingsteps)
```

Impute missing values in the dataset with the average for that interval  
The average is truncated to an integer
```{r}
imputedsteps <- sapply(data[missingsteps,interval], function(i) {
  as.integer(stepsperinterval[interval==i,steps])
})
```

Create a new dataset that is equal to the original dataset but with the missing
data filled in
```{r}
newdata <- copy(data)
newdata[missingsteps, steps:=imputedsteps]
```

Make a histogram of the total number of steps taken each day  and report the
new mean and median.
```{r}
newstepsperday <- newdata[, .(steps=sum(steps)), by=date]
par(mfcol=c(1,2))
with(stepsperday, hist(steps, main="Old Histogram", col="Red", ylim=c(0,40)))
with(newstepsperday, hist(steps, main="New Histogram", col="Blue", ylim=c(0,40)))
summary(newstepsperday$steps)[c("Mean", "Median")]
```




## Are there differences in activity patterns between weekdays and weekends?
Create a new factor variable indicating whether a given date is a weekday or weekend
```{r}
weekdayweekend <- ifelse(
  weekdays(newdata$date) %in% c("Saturday", "Sunday"),
  "Weekend", "Weekday"
)
newdata[,weekday:=as.factor(weekdayweekend)]
```
Make a time series plot of the interval and the number of steps taken during
weekdays and weekends
```{r}
newstepsperinterval <- newdata[, .(steps=mean(steps)), by=.(interval, weekday)]
library(lattice)
xyplot(
  steps~interval | weekday,
  data=newstepsperinterval,
  type="l",
  layout=c(1,2)
)
```
